---
title: "Risk Score Function Description"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, let's load the simulated data and the function. Note that the function requires three packages: caret, glmnet, and RcppEigen. They will automatically load with the "source" command but you will need to install them if you have never used them before.

```{r load}
load("SimulatedData.RData")
source("RiskScoreFunction.R")

```

The data frame columns need to be in the following order: 1 traffic-related air pollution (TRAP) exposure, covariates, metabolites. The number of covariates can vary (the function default is 5), and the number of metabolites does not matter. But the function assumes that there is only one exposure and it is in the first column, followed by the covariates and then metabolites.

The simulated data has one TRAP variable, one covariate, and 10,000 metabolites.

```{r cols}
dim(sim.met.data)
head(names(sim.met.data))
```

Function input (in order):

 - data frame (no default)
 - Number of covariates (default is 5)
 - vector containing p-value thresholds (no default)
 - FDR threshold for Benjamini-Hochberg correction (default is 0.2)
 - logical indicator for missing data (default is MISS=FALSE). When MISS=TRUE, the function will skip elastic net
 - seed (no default) can be any  number, but helps with reproducibility (e.g., can use three different seeds to compare results of three different train/test splits)
 - level to indicate desired confidence level (0.95 default)
 
The output is a set of lists, one for each p-value threshold, the Benjamini-Hochberg correction, and one for elastic net, if applicable. Each list contains the following:

 - captured: the number of metabolites captured
 - RiskScores: vector of risk scores
 - Coefficients: a matrix containing the coefficients, SEs, p-values, and confidence intervals
 - Rsquare: the R square value
 - MetNames: column names for the metabolites used in the risk score
 - RSCoefs: Coefficient (weight) values used in the risk score calculation
 
Example function call and output:

```{r example}
myseed <- 123
pval.vec <- c(.01,.001,.0001,.00001,.000001)
myres <- RiskScoreFun(DF=sim.met.data,
                      Ncov=1,
                      pval.thres.vec=pval.vec,
                      FDR=0.2,
                      miss=FALSE,
                      iseed=myseed,
                      level=0.95)

```

Now you can access the output for a particular p-value threshold, elastic net, or Benjamini-Hochberg. The lists are named by the p-value threshold ("Enet" for elastic net or "BH" for Benjamini-Hochberg)

The example below shows that using the Benjamini-Hochberg correction for FDR threshold of 0.2, 55 metabolites are selected for the risk score, and the risk score shows a significant association with the TRAP ($p<0.0001$)

```{r example2}
myres$BH$captured

myres$BH$Coefficients

```

We can use the test data to plot the risk scores and the TRAP

```{r example3}
plot(myres$BH$RiskScores,myres$test.DF$TRAP,
     xlab="Risk Scores",
     ylab="Air pollutant exposure",
     main="Visualizing the association between the risk score and the TRAP")
```